---
title: "Histone acetylation after 1 hr endurance exercise"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

Six male participants underwent 1 hr endurance exercise (~70% VO2 peak) with skeletal muscle biopsy before and immediately after exercise.
These skeletal muscle samples underwent mRNA-seq, and ChIP-seq for H3K9/K14ac and H3K36ac. 
The goal is to understand pathway regulation at the level of RNA expression and histone acetylation.
Gene expression was quantified using STAR uniquely mapped reads and promoter based histone acetylation changes were quantified by read counting within 3 kbp of the TSS.
In this analysis I will perform a multivariate analysis of REACTOME pathway enrichment.

## Obtain REACTOME gene sets

```{r,genesets}
library(mitch)
download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip", destfile="ReactomePathways.gmt.zip")
unzip("ReactomePathways.gmt.zip")
genesets <- gmt_import("ReactomePathways.gmt")
head(genesets,3)
```

## Import profiling data

```{r,import}
rna <- read.table("data/RNA.tsv",header=TRUE,row.names = 1)
head(rna)

k9a <- read.table("data/K9ac.tsv",header=TRUE,row.names = 1)
head(k9a)

k36a <- read.table("data/K36ac.tsv",header=TRUE,row.names = 1)
head(k36a)

genenames <- as.data.frame(unique(c(rownames(rna),rownames(k9a),rownames(k36a))),stringsAsFactors = FALSE)
genenames$genesymbol <- sapply(strsplit(genenames[,1],"_"),"[[",2)
colnames(genenames) <- c("GeneID","GeneName")
head(genenames)

x <- list("rna"=rna,"k9a"=k9a,"k36a"=k36a)
xx <- mitch_import(x,DEtype = "edger", geneTable = genenames)

```